---
title: "Final Project 2:R Markdown"
author: "Marian Mumin"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction - the what and why

This dataset includes a variety of demographic and health-related variables, such as stress level, life satisfaction, immigration status, nativity, self-reported health satisfaction, and behavioral indicators like drug use.

The purpose of my analysis is to examine how immigration status (identifying as an immigrant) related to perceived health outcomes as well as compares to the percieved health outcomes of those classified under nativity. 

Research questions: 
1. Explore the distribution of health satisfaction across immigrant and non-immigrant groups. 
2. investigate the relationship between immigration status and drug use/abuse, compare foreign and native-born individuals. 

My goal is to understand patterns of health satisfaction and risk behaviors among immigrants versus non-immigrants. 

```{r, laod-data, echo = FALSE, message=FALSE, warning=FALSE}

# Downloading packages necessary for codes 

library(tidyverse)
library(ggplot2)
library(knitr)
library(dplyr)
library(gtsummary)
library(gt)
library(kableExtra)

#Reading in my data
health_data <- read.csv("data/health_dataset.csv")

```

```{r birth_country_recode, echo = FALSE}

# For future reference incase i want to use birth_country in general or as a comparison for proper birth place identification
health_data <- health_data %>%
  mutate(
    Birth_country = factor(
      Birth_country,
      levels = c(1, 2, 3),
      labels = c("Canada", "Outside Canada", "Not stated")
    )
  )

```

# Table: Summary of Health Satisfaction by Immigration Status
## Tabular Analysis

The table below shows the distribution of self-reported life satisfaction scores among immigrants and native-born respondents.It reveals that "Community Belonging" is statistically significant (p < 0.001), with immigrants reporting a slightly higher "Very strong" belonging (23%) compared to Canadian-born respondents (22%). Interestingly, while the mean Life Satisfaction score is identical at 8.06 for both groups, the p-value of 0.012 suggests that the underlying distribution of these scores differs significantly between immigrants and native-born individuals. This indicates that while the "average" experience is similar, the spread or frequency of specific scores varies across the two populations.  

```{r, table1, echo = FALSE, message=FALSE, warning=FALSE}

# Table analyzing Life Satisfaction and sense of belonging by immigrant status (Canadian Born vs Landed immigrant) 

table1 <- readRDS( 
 here::here("output/table1.rds") 
  )
table1
```

## Graphical Analysis 

The figure below visualizes the distribution of life satisfaction scores for immigrants and native-born respondents. I used a boxplot to visualize those comparisons.my Graph 1 shows that the median life satisfaction sits at the same level for both groups. However, the whiskers and potential outliers likely highlight the slight distributional differences noted in the Wilcoxon rank-sum test from the table. By using a boxplot, we can see that despite the stresses of immigration, the interquartile range for satisfaction remains remarkably stable and comparable to those born in Canada. 

```{r sense of belonging and life-satisfaction figure, fig.align="center", out.width="600px", echo=FALSE, message=FALSE, warning=FALSE}

Figure1 <- readRDS( 
 here::here("output/Figure1.rds") 
  )
Figure1
```

## Graph 2: bar graph to display general health stutus between immigrants and canadian born participants 

The bar graph provides clear evidence of the "Healthy Immigrant Effect," as immigrants report "Excellent" health at a higher rate (23.1%) than Canadian-born participants (19%). Furthermore, Canadian-born respondents show higher percentages in the "Fair" and "Poor" categories, totaling 15.3% compared to only 12.5% for immigrants. These results directly answer the first research question, showing that while both groups are generally healthy, the immigrant group maintains a slight but distinct advantage in self-reported physical well-being.

```{r general_health_plot, fig.align="center", out.width="600px", echo=FALSE, message=FALSE, warning=FALSE} 
# Second graph 

Figure2 <- readRDS( 
 here::here("output/Figure2.rds") 
  )
Figure2
```
 


